% ============================================================
%  Modeling Radiotrophic Fitness
%  Hunter Kinder & Brett Faulkner
%  Preprint — bioRxiv Systems Biology — April 2026
% ============================================================
\documentclass[11pt]{article}

% --- Geometry -----------------------------------------------
\usepackage[top=1in,bottom=1in,left=1.1in,right=1.1in]{geometry}

% --- Encoding & fonts ----------------------------------------
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}

% --- Math ---------------------------------------------------
\usepackage{amsmath,amssymb,amsthm}
\usepackage{bm}

% --- Tables -------------------------------------------------
\usepackage{booktabs}
\usepackage{array}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{rotating}
\usepackage{pdflscape}

% --- Graphics & colour ----------------------------------------
\usepackage{xcolor}
\definecolor{linkblue}{RGB}{0,70,150}

% --- Hyperref ------------------------------------------------
\usepackage[colorlinks=true,
            linkcolor=linkblue,
            citecolor=linkblue,
            urlcolor=linkblue,
            pdftitle={Modeling Radiotrophic Fitness},
            pdfauthor={Hunter Kinder, Brett Faulkner}]{hyperref}

% --- Layout ---------------------------------------------------
\usepackage{parskip}
\setlength{\parskip}{6pt}
\usepackage{setspace}
\setstretch{1.15}
\usepackage{titlesec}
\titleformat{\section}{\large\bfseries}{\thesection.}{0.5em}{}
\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection}{0.5em}{}
\titlespacing{\section}{0pt}{14pt}{6pt}
\titlespacing{\subsection}{0pt}{10pt}{4pt}

% --- Floats & captions ----------------------------------------
\usepackage{caption}
\captionsetup{font=small,labelfont=bf}

% (no datetime2 in basic install)

% ============================================================
\begin{document}

% ---- Title block --------------------------------------------
\begin{center}
  {\LARGE\bfseries Modeling Radiotrophic Fitness}\\[10pt]
  {\large\bfseries Spatiotemporal Dynamics of Microbial Communities\\
   Under Ionizing Radiation Stress}\\[14pt]
  {\normalsize
    \textbf{Hunter Kinder}, B.A., M.A.T.L. --- Independent Researcher, Missouri, USA\\[2pt]
    \textbf{Brett Faulkner}, B.Sc, BioPhys --- Independent Researcher\\[10pt]
  }
  {\small
    \textit{Preprint --- submitted to bioRxiv (Systems Biology)}\\
    \textit{Version 1.0 --- April 2026}
  }
\end{center}
\vspace{4pt}
\noindent\rule{\linewidth}{0.5pt}
\vspace{2pt}

% ---- Abstract ------------------------------------------------
\begin{abstract}
\noindent
This work presents a mathematical framework to describe the radiotrophic and adaptive dynamics
of microbial communities exposed to gamma radiation. The developed model integrates Langevin
dynamics with reaction-diffusion equations to simulate the spatial motility, growth, and nonlinear
interactions of radiotrophic fungi, extremophiles, and radiation-sensitive averse species. In this
framework, radiation gradients and chemical factors, such as melanin synthesis, are introduced as
modifiable parameters to analyze their influence on biofilm growth and adaptive responses. The
model leverages stochastic differential equations to capture complex dependencies between
environmental conditions and microbial interactions. Numerical simulations of seven interacting
species---including \textit{Cryptococcus neoformans}, \textit{Deinococcus radiodurans},
\textit{Cladosporium sphaerospermum}, \textit{Bacillus subtilis}, \textit{Aspergillus niger},
\textit{Shewanella oneidensis}, and \textit{Ochrobactrum intermedium} AM7---demonstrate that
radiation gradients drive spatial niche partitioning, with melanized radiotrophic fungi
concentrating in high-radiation zones while radiosensitive metal-reducing bacteria occupy
attenuated peripheries. This work contributes a predictive modeling tool capable of elucidating
the nonlinear adaptive behaviors of radiotrophic communities in extreme environments, with
applications to nuclear bioremediation.
\end{abstract}
\noindent\rule{\linewidth}{0.5pt}

% ---- 1. Introduction -----------------------------------------
\section{Introduction}

Microbial communities possess an extraordinary ability to adapt and survive under extreme
environmental stressors, such as elevated radiation levels. Within such communities, radiotrophic
fungi and other extremophilic microorganisms have evolved specialized mechanisms that allow them
to thrive in high-radiation environments. This study aims to develop a comprehensive mathematical
framework that captures the complex interplay between these organisms' nonlinear interactions,
radiation sensitivities, and chemotactic behaviors within biofilm communities. By leveraging a
combined approach that integrates Langevin dynamics, reaction-diffusion modeling, and a
Hamiltonian-based formulation, we simulate and explore the spatiotemporal evolution of these
communities under radiation stress and other external influences.

The practical motivation for this work lies in nuclear bioremediation. Contaminated sites such
as the Chernobyl Exclusion Zone and the Hanford Nuclear Reservation host complex microbial
communities in which radiotrophic species coexist with metal-reducing bacteria capable of
immobilizing radionuclides. Mathematical models that predict community dynamics under spatially
heterogeneous radiation fields are essential for designing and optimizing bioremediation
strategies. The framework developed here provides the theoretical foundation for such predictions.

% ---- 2. Related Work -----------------------------------------
\section{Related Work}

\subsection{Radiotrophic Organisms and Melanin-Mediated Radiotrophy}

The discovery that melanized fungi thrive in high-radiation environments has fundamentally
reshaped our understanding of biological energy transduction. Zhdanova et al.\ \cite{zhdanova2000}
first documented the prevalence of darkly pigmented fungi within the damaged reactor at the
Chernobyl Nuclear Power Plant, noting that species such as \textit{Cladosporium sphaerospermum}
dominated the mycobiota of the most contaminated structures. Subsequent work by Dadachova et
al.\ \cite{dadachova2007} demonstrated that ionizing radiation alters the electronic properties of
melanin, enhancing electron-transfer activity in melanized cells of \textit{Cryptococcus neoformans}
and \textit{Wangiella dermatitidis}. Melanized cells exposed to radiation levels approximately 500
times above background grew significantly faster, accumulated more biomass, and incorporated
three-fold more ${}^{14}$C-acetate than non-irradiated melanized cells or irradiated albino
mutants. Dadachova and Casadevall \cite{dadachova2008} subsequently proposed the term
``radiosynthesis'' to describe this phenomenon, drawing an analogy to photosynthesis in which
melanin serves a role loosely comparable to chlorophyll. More recently, Shunk et al.\
\cite{shunk2022} cultivated \textit{C.\ sphaerospermum} aboard the International Space Station
for 26 days, demonstrating a growth advantage of approximately 21\% under space radiation
conditions and measurable attenuation of ionizing radiation beneath the fungal biomass.

The radiation resistance of \textit{Deinococcus radiodurans} operates through fundamentally
different mechanisms. Rather than harnessing radiation for metabolic benefit, \textit{D.\
radiodurans} survives doses exceeding 12~kGy through a combination of efficient DNA
double-strand break repair via extended synthesis-dependent strand annealing (ESDSA), multiple
genome copies enabling homologous recombination, and a potent manganese-based antioxidant system
that protects proteins from oxidative damage \cite{slade2011,daly2009}. Daly et al.\
\cite{daly2010} showed that small-molecule Mn$^{2+}$-metabolite complexes specifically protect
the proteome against radiation-induced carbonylation, establishing that protein protection, rather
than DNA repair capacity alone, governs extreme radioresistance. \textit{Aspergillus niger},
while not radiotrophic in the strict sense, produces melanin that confers substantial
radioprotection, with melanized fungi generally exhibiting LD$_{10}$ values approaching or
exceeding 1~kGy \cite{dadachova2008}.

\subsection{Mathematical Models of Biofilm Communities}

The mathematical modeling of biofilm communities has a rich history beginning with the
foundational one-dimensional multispecies model of Wanner and Gujer \cite{wanner1986}, which
coupled reaction-diffusion equations for substrate transport with biomass conservation laws to
predict biofilm thickness and spatial species distributions. Xavier et al.\ \cite{xavier2005}
extended this framework to multiple spatial dimensions, providing a deterministic continuum
approach to heterogeneous biofilm development. Individual-based modeling approaches, exemplified
by the iDynoMiCS platform of Lardon et al.\ \cite{lardon2011}, complement continuum models by
resolving cell-level stochastic interactions. Cross-diffusion biofilm models have been analyzed
by Sonner et al.\ \cite{sonner2015}, who established mathematical properties of volume-filling
multispecies systems that exhibit porous-medium-type degeneracy. Our PSDE framework extends
these precedents by incorporating radiation-dependent fitness terms and phase-locking dynamics
that couple species interactions to external radiation fields, moving beyond the
substrate-limited growth paradigms that dominate existing biofilm models.

\subsection{Hamiltonian and Stochastic Approaches in Theoretical Ecology}

The application of Hamiltonian mechanics to ecological systems remains comparatively
underexplored, despite the natural conservation laws that govern energy flow in ecosystems.
Symplectic integration methods, which preserve the geometric structure of Hamiltonian flows,
have been advocated for ecological modeling by Diele et al.\ \cite{diele2015} to ensure that
numerical solutions maintain essential qualitative dynamics. The Kuramoto model of coupled phase
oscillators \cite{kuramoto1984} provides a well-established framework for synchronization
phenomena in biological systems, from neural networks to circadian rhythms. Aceb\'r\'on et al.\
\cite{acebron2005} reviewed the model as a paradigm for synchronization, noting its
generalization to systems with time delays, frequency-weighted coupling, and heterogeneous
interaction topologies. Our phase-locking kernel $\Gamma_s(t,\mathbf{x})$ draws directly on
this tradition, extending it to multispecies microbial communities where species-level
oscillatory dynamics in growth and resource acquisition couple through shared radiation and
nutrient fields. The use of Langevin stochastic dynamics to describe microbial population
fluctuations follows naturally from the Fokker-Planck formalism and has been applied to
competing microbial populations in chemostat systems \cite{campillo2017}, where stochastic
differential equations revealed long-term competitive outcomes differing from deterministic
predictions.

\subsection{Bioremediation Context}

The practical motivation for modeling radiotrophic communities lies in nuclear bioremediation.
\textit{Shewanella oneidensis} MR-1 is a model dissimilatory metal-reducing bacterium capable
of reducing soluble U(VI) to insoluble U(IV) via extracellular electron transfer through
c-type cytochromes \cite{heidelberg2002,veeramani2011}. \textit{Ochrobactrum intermedium} AM7,
isolated from soil near the Kakrapar Atomic Power Station in India, tolerates high concentrations
of thorium and produces exopolysaccharides (EPS) that complex with Th(IV), offering a pathway
toward bioremediation of actinide-contaminated environments \cite{shukla2020}. The combination
of radiotrophic fungi that actively benefit from radiation with metal-reducing bacteria that
immobilize radionuclides suggests the possibility of engineered biofilm consortia for in situ
remediation. The fitness landscape framework introduced by Wright \cite{wright1932} and
formalized through the NK model of Kauffman \cite{kauffman1989} provides a conceptual foundation
for understanding how epistatic interactions among species traits shape adaptation under
radiation stress, a perspective that our Hamiltonian kNN decision tree operationalizes
quantitatively.

% ---- 3. Mathematical Framework --------------------------------
\section{Mathematical Framework}

\subsection{Symbols and Notation}

\begin{table}[h!]
\centering
\small
\caption{Summary of principal mathematical symbols.}
\label{tab:notation}
\begin{tabular}{@{}lp{0.72\linewidth}@{}}
\toprule
\textbf{Symbol} & \textbf{Description} \\
\midrule
$F_s(t,\mathbf{x})$ & Fitness function of species $s$, evolving over time $t$ and space $\mathbf{x}$ \\
$H(p,q)$            & Hamiltonian function defining total system energy \\
$p,\,q$             & Phase-space variables (generalised momentum and position) \\
$\eta_s(t,\mathbf{x}) = \sigma_s\xi(t,\mathbf{x})$ & Stochastic noise: $\sigma_s$ intensity; $\xi$ spatiotemporal white noise \\
$D_s$               & Species-specific passive diffusion coefficient \\
$\mu_s$             & Species-specific directed-motility (chemotaxis) coefficient \\
$P_{sj}(t)$         & Permutation matrix describing species-$s$ to species-$j$ transitions \\
$I_\gamma(t,\mathbf{x})$ & Ionizing (gamma) radiation field \\
$N(t,\mathbf{x})$   & Non-ionizing (UV) radiation field \\
$k_B$               & Boltzmann constant \\
$S[q(t)]$           & Action functional for species trajectories \\
$\nabla$            & Spatial gradient operator \\
$\Gamma_s(t,\mathbf{x})$ & Phase-locked kernel for species $s$ \\
$\beta_{s,\mathrm{ion}}$ & Ionizing radiation sensitivity coefficient \\
$\alpha_{s,\mathrm{nir}}$ & Non-ionizing radiation coupling coefficient \\
$\theta_s$          & Adaptive phase-locking coefficient \\
$\gamma_s$          & Positive gamma-radiation sensitivity (benefit) coefficient \\
$\Delta_s$          & Phase-locking synchronization deviation \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Partial Stochastic Differential Equation (PSDE) for Species Fitness}

The fitness of each species $s$ is governed by the following PSDE:
\begin{equation}
\label{eq:psde}
\partial_t F_s = \underbrace{\nabla\!\cdot(D_s\nabla F_s)}_{\text{diffusion}}
  - \underbrace{\nabla\!\cdot\!\left(\mu_s\sum_{j=1}^n P_{sj}(t)\,F_j\right)}_{\text{directed motility}}
  + \underbrace{R_s}_{\text{growth}}
  + \underbrace{\sigma_s\xi(t,\mathbf{x})}_{\text{noise}}
  - \underbrace{\beta_{s,\mathrm{ion}}\,I_\gamma F_s}_{\text{ionising damage}}
  + \underbrace{\gamma_s\Delta_s}_{\text{phase-lock gain}}
  - \underbrace{\alpha_{s,\mathrm{nir}}\,N F_s}_{\text{UV coupling}}
  + \underbrace{\theta_s H_s}_{\text{Hamiltonian interaction}}
  + \underbrace{C_s}_{\text{external forcing}}
\end{equation}

\noindent where $F_s \equiv F_s(t,\mathbf{x})$, $I_\gamma \equiv I_\gamma(t,\mathbf{x})$, and
$N \equiv N(t,\mathbf{x})$.  The nine terms encode:
\begin{itemize}\setlength\itemsep{2pt}
  \item \textbf{Diffusion}: passive spatial spreading at rate $D_s$.
  \item \textbf{Directed motility}: chemotaxis and radiation-gradient-directed movement
        controlled by $\mu_s$ and the transition matrix $P_{sj}$.
  \item \textbf{Deterministic reaction}: $R_s(t,\mathbf{x})$, nutrient-dependent growth
        (see Section~\ref{sec:nutrient}).
  \item \textbf{Stochastic noise}: $\sigma_s\xi(t,\mathbf{x})$, spatiotemporal white noise
        scaled by species noise intensity $\sigma_s$.
  \item \textbf{Ionising damage}: $-\beta_{s,\mathrm{ion}}\,I_\gamma(t,\mathbf{x})\,F_s$,
        radiation-induced fitness reduction.
  \item \textbf{Phase-locking adjustment}: $\gamma_s\Delta_s$, adaptive synchronisation gain
        from melanin-mediated energy transduction.
  \item \textbf{Non-ionising effect}: $-\alpha_{s,\mathrm{nir}}\,N(t,\mathbf{x})\,F_s$,
        UV/heat-driven fitness coupling.
  \item \textbf{Hamiltonian interaction}: $\theta_s H_s(t,\mathbf{x})$, energy-conserving
        inter-species force from the Hamiltonian (Section~\ref{sec:hamiltonian}).
  \item \textbf{External forcing}: $C_s(t,\mathbf{x})$, environmental control inputs.
\end{itemize}

\subsection{Hamiltonian Framework}
\label{sec:hamiltonian}

The total Hamiltonian for the multi-species biofilm is:
\begin{equation}
\label{eq:hamiltonian}
H = \sum_{i=1}^{N}\!\left[\frac{1}{2}\,\rho_i\,v_i^2 + U_i(\mathbf{x}_i)\right]
  + \sum_{i\neq j} V_{ij}(r_{ij})
  + \sum_{k=1}^{K} W_k(t,\mathbf{x})
\end{equation}

\noindent where $\rho_i$ is the effective biomass density of species $i$ [cells\,$\mu$m$^{-3}$
$\times$ effective mass analog], $U_i(\mathbf{x}_i)$ is the potential energy in the nutrient
concentration field, $V_{ij}(r_{ij})$ is the pairwise interaction energy, and $W_k(t,\mathbf{x})$
captures external energy contributions from radiation exposure and stochastic effects. For
mutualistic interactions:
\begin{equation}
  V_{ij}^{\mathrm{mutual}}(r_{ij}) = -\gamma\,\exp\!\left(-\frac{r_{ij}^2}{\sigma^2}\right)
\end{equation}

\subsection{Symplectic Integration}

Symplectic integrators preserve the geometric structure of the Hamiltonian flow, ensuring that
energy and momentum do not drift over extended simulation windows. Phase-space trajectories
evolve via the phase-locked canonical equations:
\begin{equation}
\label{eq:canonical}
\frac{dq}{dt} = \frac{\partial H}{\partial p} - \Gamma_s(t,\mathbf{x})\,F_s(t,\mathbf{x}),
\qquad
\frac{dp}{dt} = -\frac{\partial H}{\partial q} + \Gamma_s(t,\mathbf{x})\,F_s(t,\mathbf{x})
\end{equation}

The leapfrog/Verlet scheme is employed numerically: (1) evaluate $H_s$ from the current state;
(2) adjust the phase-locking deviation $\Delta_s$; (3) advance via a symplectic step
\cite{hairer2006}.

\subsection{Radiation Field Models}

\paragraph{Ionizing radiation (gamma decay).}
\begin{equation}
\label{eq:gamma_field}
I_\gamma(t,\mathbf{x}) = I_0\,\exp(-\kappa\,\|\mathbf{x}\|)
\end{equation}
where $I_0$ is the initial intensity at the source and $\kappa$ is the attenuation coefficient
scaled by local biofilm density.

\paragraph{Non-ionizing radiation (UV oscillatory field).}
\begin{equation}
\label{eq:uv_field}
N(t,\mathbf{x}) = N_{\mathrm{UV}}\cos(\omega t)
\end{equation}

\subsection{Melanin Reaction-Diffusion}

The spatial production and diffusion of melanin by radiotrophic fungi is modelled as:
\begin{equation}
\label{eq:melanin}
\frac{\partial M}{\partial t} = D_M\,\nabla^2 M + \alpha_M\cdot n_{\mathrm{RF}}(\mathbf{x},t)\cdot I_\gamma(t,\mathbf{x})
\end{equation}
where $D_M$ is the melanin diffusion coefficient [$\mu$m$^2$\,s$^{-1}$], $\alpha_M$ is the
radiation-driven production rate [$\mu$g cell$^{-1}$ Gy$^{-1}$], and $n_{\mathrm{RF}}$ is the
local radiotrophic cell density [cells\,$\mu$m$^{-3}$]. Each term carries units of
[$\mu$g\,$\mu$m$^{-3}$\,s$^{-1}$].

\subsection{Nutrient Uptake with Adaptive Feedback}
\label{sec:nutrient}

\begin{equation}
\label{eq:nutrient}
R_s(t,\mathbf{x}) = \alpha_s\,C_s(t,\mathbf{x})\!\left(1 - \frac{F_s(t,\mathbf{x})}{K_s}\right)
  + A_s(t)\,C_s(t,\mathbf{x})
\end{equation}
\begin{equation}
\label{eq:adaptive}
A_s(t) = \gamma_s\,\phi_s(t) - \beta_{s,\mathrm{ion}}\,I_\gamma(t)
\end{equation}
where $\phi_s(t) = \cos(\omega_s t - \theta_s)$ is the phase-alignment function and $A_s(t)$ is
the adaptive feedback driven by phase-locking and local radiation stress.

\subsection{Hamiltonian kNN Decision Tree}

To model species transition probabilities modulated by neighbourhood phase states:
\begin{equation}
\label{eq:knn}
H_{\mathrm{kNN}} = \frac{1}{\sigma_n}\sum_{j=1}^{n} P_{sj}(t)\,F_j(t,\mathbf{x})\,\Gamma_s(t,\mathbf{x})
\end{equation}

\subsection{Radiation Damage and Resilience}

\begin{equation}
\label{eq:damage}
D_{\mathrm{rad},s}(t,\mathbf{x}) = -\beta_{s,\mathrm{rad}}\,I_\gamma(t,\mathbf{x})\,F_s(t,\mathbf{x})
\end{equation}
Species with higher melanin production or stronger DNA repair exhibit lower $\beta_{s,\mathrm{rad}}$
values, reflecting the Mn$^{2+}$-dependent proteome protection described by Daly et al.\
\cite{daly2010}.

\subsection{Biofilm Mechanical Properties}

The EPS matrix is modelled as a Finite Extensible Nonlinear Elastic (FENE) entropic spring:
\begin{equation}
\label{eq:fene}
U(r) = -\frac{1}{2}\,k\,R^2\,\ln\!\left(1 - \frac{r^2}{R^2}\right)
\end{equation}
where $k$ is the spring constant, $R$ is the maximum extensibility, and $r$ is the current
end-to-end distance.  Viscoelastic creep follows the Kelvin-Voigt model:
\begin{equation}
\label{eq:kelvinvoigt}
\sigma(t) = E\,\varepsilon(t) + \eta\,\frac{d\varepsilon}{dt}
\end{equation}
with stress relaxation time $\tau = \eta/E$ governing biofilm restructuring dynamics.

% ---- 4. Parameter Estimation ----------------------------------
\section{Parameter Estimation}

Biologically justified parameter ranges for the six primary modelled species, derived from
published experimental data, are given in Table~\ref{tab:params}.

\begin{landscape}
\begin{table}[p]
\centering
\footnotesize
\caption{Biologically justified parameter ranges for modelled species.  Values derived from
         published experimental data.}
\label{tab:params}
\setlength\tabcolsep{3.5pt}
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{@{}lllllp{6cm}c@{}}
\toprule
\textbf{Parameter} & \textbf{Symbol} & \textbf{Species} & \textbf{Range} & \textbf{Units} & \textbf{Biological Basis} & \textbf{Ref} \\
\midrule
Diffusion coeff. & $D_s$ & \textit{C.\ neoformans}     & $0.01$--$0.10$            & $\mu$m$^2$ s$^{-1}$ & Passive Brownian diffusion, $\sim$5\,$\mu$m yeast & \cite{berg1993} \\
Diffusion coeff. & $D_s$ & \textit{D.\ radiodurans}    & $0.05$--$0.50$            & $\mu$m$^2$ s$^{-1}$ & Small coccoid ($\sim$1.5\,$\mu$m); higher $D$ from size & \cite{slade2011} \\
Diffusion coeff. & $D_s$ & \textit{B.\ subtilis}       & $0.10$--$1.00$            & $\mu$m$^2$ s$^{-1}$ & Rod-shaped motile cells; active motility augments $D$ & \cite{guttenplan2013} \\
Diffusion coeff. & $D_s$ & \textit{C.\ sphaerospermum} & $0.005$--$0.05$           & $\mu$m$^2$ s$^{-1}$ & Filamentous hyphae; low effective diffusion & \cite{shunk2022} \\
Diffusion coeff. & $D_s$ & \textit{A.\ niger}          & $0.005$--$0.05$           & $\mu$m$^2$ s$^{-1}$ & Filamentous; comparable to \textit{C.\ sphaerospermum} & \cite{dadachova2008} \\
Diffusion coeff. & $D_s$ & \textit{S.\ oneidensis}     & $0.10$--$0.80$            & $\mu$m$^2$ s$^{-1}$ & Facultative anaerobe, rod-shaped, flagellated & \cite{heidelberg2002} \\
\midrule
Motility coeff. & $\mu_s$ & \textit{C.\ neoformans}    & $0.00$--$0.05$  & $\mu$m s$^{-1}$ & Non-motile yeast; residual drift from EPS flow & \cite{dadachova2007} \\
Motility coeff. & $\mu_s$ & \textit{D.\ radiodurans}   & $0.00$--$0.01$  & $\mu$m s$^{-1}$ & Non-motile coccus & \cite{slade2011} \\
Motility coeff. & $\mu_s$ & \textit{B.\ subtilis}      & $15$--$45$      & $\mu$m s$^{-1}$ & Flagellum-driven; $\sim$25\,$\mu$m s$^{-1}$ at 30\,$^\circ$C & \cite{guttenplan2013} \\
Motility coeff. & $\mu_s$ & \textit{C.\ sphaerospermum}& $0.00$--$0.005$ & $\mu$m s$^{-1}$ & Hyphal extension; 0.1--5\,$\mu$m min$^{-1}$ tip growth & \cite{shunk2022} \\
Motility coeff. & $\mu_s$ & \textit{S.\ oneidensis}    & $10$--$40$      & $\mu$m s$^{-1}$ & Polar flagellum; comparable to \textit{E.\ coli} & \cite{heidelberg2002} \\
\midrule
Ion.\ rad.\ sens. & $\beta_{s,\mathrm{ion}}$ & \textit{C.\ neoformans}     & $10^{-5}$--$10^{-4}$              & Gy$^{-1}$ & Melanised; LD$_{10}$ $\sim$2--5\,kGy & \cite{dadachova2007,dadachova2008} \\
Ion.\ rad.\ sens. & $\beta_{s,\mathrm{ion}}$ & \textit{D.\ radiodurans}    & $5\!\times\!10^{-6}$--$5\!\times\!10^{-5}$ & Gy$^{-1}$ & $D_{10}$ $\sim$12\,kGy; most radioresistant organism known & \cite{slade2011,daly2009} \\
Ion.\ rad.\ sens. & $\beta_{s,\mathrm{ion}}$ & \textit{B.\ subtilis}       & $10^{-3}$--$5\!\times\!10^{-3}$  & Gy$^{-1}$ & Spore $D_{10}$ $\sim$1--2\,kGy; vegetative $\sim$0.2--0.6\,kGy & \cite{nicholson2000} \\
Ion.\ rad.\ sens. & $\beta_{s,\mathrm{ion}}$ & \textit{C.\ sphaerospermum} & $10^{-5}$--$10^{-4}$              & Gy$^{-1}$ & Melanised Chernobyl isolate; $\approx$ \textit{C.\ neoformans} & \cite{zhdanova2000,shunk2022} \\
Ion.\ rad.\ sens. & $\beta_{s,\mathrm{ion}}$ & \textit{A.\ niger}          & $5\!\times\!10^{-5}$--$5\!\times\!10^{-4}$ & Gy$^{-1}$ & Melanised; less resistant than Chernobyl isolates & \cite{dadachova2008} \\
Ion.\ rad.\ sens. & $\beta_{s,\mathrm{ion}}$ & \textit{S.\ oneidensis}     & $5\!\times\!10^{-2}$--$10^{-1}$  & Gy$^{-1}$ & $D_{10}$ $\sim$0.07\,kGy; extremely radiosensitive & \cite{daly2009} \\
\midrule
Melanin rate & $\alpha_M$ & \textit{C.\ neoformans}     & $0.05$--$0.15$ & $\mu$g cell$^{-1}$ Gy$^{-1}$ & Substrate-dependent; enhanced by radiation & \cite{dadachova2007} \\
Melanin rate & $\alpha_M$ & \textit{C.\ sphaerospermum} & $0.08$--$0.20$ & $\mu$g cell$^{-1}$ Gy$^{-1}$ & Constitutively melanised; high production rate & \cite{zhdanova2000,shunk2022} \\
Melanin rate & $\alpha_M$ & \textit{A.\ niger}          & $0.03$--$0.10$ & $\mu$g cell$^{-1}$ Gy$^{-1}$ & DHN-melanin pathway & \cite{dadachova2008} \\
\midrule
Carrying cap. & $K_s$ & \textit{C.\ neoformans}    & $10^4$--$10^6$ & cells mm$^{-3}$ & Yeast biofilm under nutrient-rich conditions & \cite{dadachova2007} \\
Carrying cap. & $K_s$ & \textit{D.\ radiodurans}   & $10^5$--$10^7$ & cells mm$^{-3}$ & Dense tetrads; high packing efficiency & \cite{slade2011} \\
Carrying cap. & $K_s$ & \textit{B.\ subtilis}      & $10^5$--$10^7$ & cells mm$^{-3}$ & Dense biofilm with EPS matrix & \cite{guttenplan2013} \\
Carrying cap. & $K_s$ & \textit{S.\ oneidensis}    & $10^5$--$10^7$ & cells mm$^{-3}$ & Planktonic and biofilm modes & \cite{heidelberg2002} \\
\midrule
Phase-lock freq. & $\omega_s$ & All species & $0.01$--$1.0$   & rad hr$^{-1}$ & Circadian/ultradian metabolic oscillation & \cite{kuramoto1984,acebron2005} \\
Noise intensity  & $\sigma_s$ & All species & $0.001$--$0.05$ & ---           & Thermal and demographic stochasticity & \cite{campillo2017} \\
\bottomrule
\end{tabular}
\end{table}
\end{landscape}

% ---- 5. Computational Methods --------------------------------
\section{Computational Methods}

Simulations were implemented in Julia using \texttt{DifferentialEquations.jl} for numerical
integration of the PSDE system (Eq.~\ref{eq:psde}).  Spatial species distributions were
initialised with small random perturbations around a central point and evolved on the
$[0,1]^2$ domain.  The Euler-Maruyama scheme was applied for stochastic integration;
Runge-Kutta (RK4) was used for deterministic validation runs.  Symplectic leapfrog integration
was employed for the Hamiltonian phase-space trajectories (Eq.~\ref{eq:canonical}).

Optimisation of species-specific parameters was performed with \texttt{JuMP.jl} and
\texttt{Ipopt}.  Sensitivity analysis employed Sobol indices (global variance decomposition)
and Morris screening to identify the highest-influence parameters ($\mu_s$, $D_s$,
$\beta_{s,\mathrm{ion}}$, $\theta_s$).  Visualisation used \texttt{PlotlyJS.jl} for interactive
3D plots, \texttt{Agents.jl} for agent-based validation runs, and \texttt{Plots.jl} with
\texttt{GR} backend for static trajectory outputs.  The interactive simulation includes sliders
for gamma radiation intensity, thorium intensity, and heat intensity at 45-step resolution.

% ---- 6. Results ---------------------------------------------
\section{Results}

\subsection{Species Trajectory Clustering Under Radiation Gradients}

The Langevin diffusion simulations on the $[0,1]^2$ domain reveal pronounced spatial phenotype
separation under the imposed radiation gradient (Figure~1). $k$-means clustering of species
trajectories identifies three distinct spatial niches. The first cluster, concentrated in the
high-radiation zone, is dominated by the melanised species \textit{C.\ neoformans} and
\textit{C.\ sphaerospermum}, whose positive gamma sensitivity coefficients ($\gamma_s \approx
0.07$ and $0.06$, respectively) enable them to exploit ionising radiation as a metabolic
stimulus. The second cluster occupies the intermediate-radiation zone and contains
\textit{D.\ radiodurans} and \textit{A.\ niger}---species that tolerate radiation without
deriving direct metabolic benefit. The third cluster, in the low-radiation periphery,
comprises \textit{B.\ subtilis}, \textit{S.\ oneidensis}, and marine taxa
(\textit{Pseudoalteromonas} sp., \textit{Polaribacter} sp., \textit{Flavobacterium} sp.),
whose higher radiation sensitivities confine them to regions where the ionising flux has
attenuated below their damage thresholds. This spatial segregation is consistent with
radiation-mediated niche partitioning, in which ionising radiation simultaneously serves as
stressor and resource---a dual role with no direct analogue in classical substrate-competition
biofilm models.

\subsection{Thorium-232 Decay Integration}

The Th-232 decay model (Figure~2) compares actual versus predicted radioactive decay on a
logarithmic scale spanning billions of years, incorporating stochastic perturbation around
the deterministic exponential decay law. The 14.05-billion-year half-life produces a
quasi-static radiation source on biological timescales (hours to years), but the stochastic
perturbation term captures quantum-mechanical uncertainty in individual decay events, which
manifests as Poisson-distributed dose-rate fluctuations. For biofilm modelling purposes,
Th-232 provides an effectively constant ionising background, while stochastic perturbation
introduces dose-rate heterogeneity at the spatial scale of microcolonies (tens of
micrometres), driving the phase-space fluctuations captured by the Langevin dynamics.

\subsection{Motility--Diffusion--Gamma Sensitivity Relationships}

The three-dimensional scatter relating motility ($\mu_s$), diffusion coefficient ($D_s$), and
gamma sensitivity $\gamma_s$ across species (Figure~3) reveals a striking inverse relationship
between radiation sensitivity and motility. \textit{C.\ neoformans}, at the highest gamma
sensitivity ($\approx 0.07$), exhibits effectively zero motility, consistent with the
hypothesis that radiotrophic organisms invest resources in melanin production rather than
active locomotion. \textit{D.\ radiodurans}, with the lowest $\beta_{s,\mathrm{ion}}$, is
likewise non-motile but for different mechanistic reasons: radioresistance derives from
molecular protection rather than radiation exploitation. The motile species
(\textit{B.\ subtilis}, \textit{S.\ oneidensis}) occupy the high-motility, low-$\gamma_s$
region, suggesting a trade-off between radiation exploitation and behavioural avoidance.
As radiation intensity rises, the competitive advantage shifts from motile radiation-avoiders
to sessile radiotrophic beneficiaries.

\subsection{Phase-Locking and Hamiltonian kNN Dynamics}

The Hamiltonian kNN decision tree (Figure~4) reveals that synchronised metabolic oscillations
emerge among species occupying the same spatial niche. Within the high-radiation cluster,
\textit{C.\ neoformans} and \textit{C.\ sphaerospermum} exhibit strong phase coherence in
their melanin production cycles, suggesting cooperative dynamics in which shared
melanin-derived radioprotection benefits both. Species in different radiation niches show weak
phase coupling, indicating competitive exclusion at niche boundaries. The kNN decision tree
correctly predicts species dominance transitions: as simulated radiation intensity increases
beyond 1~kGy, \textit{B.\ subtilis} populations collapse, \textit{D.\ radiodurans} maintains
steady-state abundance, and \textit{C.\ neoformans} increases in fitness---reproducing the
biologically expected ordering of radiation tolerance.

% ---- 7. Discussion -------------------------------------------
\section{Discussion}

\subsection{The Hamiltonian Framework Versus Classical Biofilm Models}

The phase-locked Hamiltonian framework differs fundamentally from prior reaction-diffusion
biofilm models in its treatment of inter-species interactions and energy conservation.
Classical models following the Wanner-Gujer tradition \cite{wanner1986} describe biofilm
dynamics through coupled PDEs governing substrate consumption and biomass growth, with species
interactions mediated exclusively through competition for shared substrates. Our Hamiltonian
$H(p,q)$ explicitly conserves total system energy across kinetic, potential, and interaction
terms, ensuring that numerical integration via the symplectic leapfrog scheme does not
introduce artificial energy drift \cite{hairer2006}. The stochastic Langevin layer adds
thermal and demographic noise while preserving the underlying Hamiltonian structure through
the fluctuation-dissipation theorem. This dual structure captures phenomena that ODE/PDE
models inherently miss: synchronised metabolic oscillations through phase-locking,
energy-conserving transitions between fitness states, and the distinction between radiation
as a damaging perturbation versus radiation as an energy source positively coupled to the
Hamiltonian.

\subsection{Biological Interpretation of \textit{C.\ neoformans} and \textit{D.\ radiodurans} Dynamics}

The simulation trajectories for \textit{C.\ neoformans} and \textit{D.\ radiodurans} align
with known biology while illuminating an underappreciated distinction between radiotrophy and
radioresistance. \textit{C.\ neoformans} in the model increases its fitness monotonically
with radiation intensity up to a species-specific threshold, consistent with the experimental
findings of Dadachova et al.\ \cite{dadachova2007} that melanised cells show enhanced
metabolic activity and faster growth under ionising radiation. The model captures the
mechanism through the positive gamma sensitivity term $\gamma_s$, which converts absorbed
dose into fitness gain via the melanin-mediated electron-transfer pathway
\cite{turick2011,casadevall2017}. \textit{D.\ radiodurans}, by contrast, maintains
approximately constant fitness across a wide dose range, declining only at extreme doses.
This flat fitness profile emerges from the balance between the radiation damage term
$\beta_{s,\mathrm{ion}}$ and the organism's molecular protection coefficient, reflecting the
Mn$^{2+}$-dependent proteome shielding identified by Daly et al.\ \cite{daly2010}. The model
thus correctly predicts that \textit{C.\ neoformans} outcompetes \textit{D.\ radiodurans} in
moderately irradiated environments where radiation provides metabolic benefit, while
\textit{D.\ radiodurans} prevails at extreme doses where even melanin-mediated protection is
overwhelmed.

\subsection{Implications for Nuclear Bioremediation}

The model's predictions for a mixed \textit{S.\ oneidensis} and \textit{O.\ intermedium} AM7
biofilm in a Th-contaminated environment suggest a synergistic remediation strategy.
\textit{S.\ oneidensis}, with its capacity for extracellular electron transfer to metal oxides
\cite{heidelberg2002,veeramani2011}, can reduce soluble actinide species and immobilise them
as insoluble precipitates. However, its extreme radiosensitivity ($D_{10} \approx 0.07$~kGy)
limits viability in high-radiation zones. The model predicts that \textit{O.\ intermedium}
AM7, with its thorium-tolerant EPS production \cite{shukla2020}, establishes a protective
biofilm matrix in the high-radiation zone that attenuates local dose rates sufficiently to
permit \textit{S.\ oneidensis} colonisation in the intermediate zone. This spatial
partitioning---emergent from the Langevin dynamics rather than imposed as a boundary
condition---suggests that self-organising biofilm architectures may naturally optimise
remediation performance. The potential for engineering \textit{D.\ radiodurans} strains for
combined radioresistance and metal reduction \cite{brim2000} offers an additional path toward
single-organism bioremediation in extreme environments.

\subsection{Mechanical Properties and Radiation Survival}

The inclusion of FENE entropic spring (Eq.~\ref{eq:fene}) and Kelvin-Voigt viscoelastic
(Eq.~\ref{eq:kelvinvoigt}) components addresses a dimension of biofilm physics largely absent
from ecological radiation models. Biofilm EPS matrices exhibit viscoelastic behaviour that
mediates mechanical stress transmission, nutrient diffusion resistance, and protection from
external perturbations. The FENE potential provides finite extensibility that prevents
unphysical deformation under radiation-induced swelling or thermal expansion, while the
Kelvin-Voigt element captures stress relaxation on timescales relevant to biofilm
restructuring. Species with higher EPS production rates (\textit{O.\ intermedium} AM7,
\textit{B.\ subtilis}) generate mechanically stiffer local environments that resist
radiation-induced structural degradation, creating protected microniches for radiosensitive
community members.

\subsection{Limitations and Future Directions}

Several simplifying assumptions limit the current model's applicability to real contaminated
sites. The radiation field is treated as spatially homogeneous with exponential attenuation
(Eq.~\ref{eq:gamma_field}), whereas actual environments at Chernobyl and Hanford exhibit
complex three-dimensional dose distributions governed by heterogeneous source geometries and
shielding materials. The UV oscillatory term (Eq.~\ref{eq:uv_field}) assumes a purely
sinusoidal diurnal cycle, neglecting atmospheric absorption spectra and seasonal variation.
The model does not incorporate horizontal gene transfer, which is known to spread radiation
resistance determinants among biofilm community members. Future work should couple the fitness
model to experimentally measured dose-rate maps from contaminated facilities, incorporate
spatially resolved nutrient transport using Darcy flow through the EPS matrix, and validate
predicted species clustering patterns against metagenomic surveys of radiotrophic biofilm
communities at nuclear sites.

% ---- 8. Conclusion -------------------------------------------
\section{Conclusion}

This work introduces a Hamiltonian-Langevin framework for modelling radiotrophic fitness in
multispecies biofilm communities exposed to ionising radiation. The PSDE fitness equation
(Eq.~\ref{eq:psde}) integrates radiation-dependent growth, melanin-mediated energy
transduction, stochastic demographic noise, and viscoelastic biofilm mechanics into a unified
formalism that conserves total system energy through symplectic integration. The central
result is that species niche partitioning under radiation gradients emerges naturally from the
phase-locked Hamiltonian dynamics, with radiotrophic fungi (\textit{C.\ neoformans},
\textit{C.\ sphaerospermum}) clustering in high-radiation zones, radioresistant bacteria
(\textit{D.\ radiodurans}) occupying intermediate regions, and radiosensitive metal-reducing
species (\textit{S.\ oneidensis}) confined to attenuated peripheries. These predictions are
consistent with known radiobiology and suggest design principles for engineered biofilm
consortia targeting nuclear bioremediation. Experimental validation through controlled
irradiation of mixed-species biofilms, combined with spatially resolved transcriptomics and
metabolomics, represents the natural next step toward translating this theoretical framework
into practical bioremediation strategies.

% ---- References ---------------------------------------------
\begin{thebibliography}{38}

\bibitem{zhdanova2000}
Zhdanova, N.N., Zakharchenko, V.A., Vember, V.V., Nakonechnaya, L.T. (2000).
Fungi from Chernobyl: mycobiota of the inner regions of the containment structures of the
damaged nuclear reactor.
\textit{Mycological Research}, 104(12), 1421--1426.
\url{https://doi.org/10.1017/S0953756200002756}

\bibitem{dadachova2007}
Dadachova, E., Bryan, R.A., Huang, X., Moadel, T., Schweitzer, A.D., Aisen, P.,
Nosanchuk, J.D., Casadevall, A. (2007).
Ionizing radiation changes the electronic properties of melanin and enhances the growth of
melanized fungi.
\textit{PLoS ONE}, 2(5), e457.
\url{https://doi.org/10.1371/journal.pone.0000457}

\bibitem{dadachova2008}
Dadachova, E., Casadevall, A. (2008).
Ionizing radiation: how fungi cope, adapt, and exploit with the help of melanin.
\textit{Current Opinion in Microbiology}, 11(6), 525--531.
\url{https://doi.org/10.1016/j.mib.2008.09.013}

\bibitem{shunk2022}
Shunk, G.K., Gomez, X.R., Kern, C., Averesch, N.J.H. (2022).
Cultivation of the dematiaceous fungus \textit{Cladosporium sphaerospermum} aboard the
International Space Station and effects of ionizing radiation.
\textit{Frontiers in Microbiology}, 13, 877625.
\url{https://doi.org/10.3389/fmicb.2022.877625}

\bibitem{slade2011}
Slade, D., Radman, M. (2011).
Oxidative stress resistance in \textit{Deinococcus radiodurans}.
\textit{Microbiology and Molecular Biology Reviews}, 75(1), 133--191.
\url{https://doi.org/10.1128/MMBR.00015-10}

\bibitem{daly2009}
Daly, M.J. (2009).
A new perspective on radiation resistance based on \textit{Deinococcus radiodurans}.
\textit{Nature Reviews Microbiology}, 7(3), 237--245.
\url{https://doi.org/10.1038/nrmicro2073}

\bibitem{daly2010}
Daly, M.J., et al. (2010).
Small-molecule antioxidant proteome-shields in \textit{Deinococcus radiodurans}.
\textit{PLoS ONE}, 5(9), e12570.
\url{https://doi.org/10.1371/journal.pone.0012570}

\bibitem{wanner1986}
Wanner, O., Gujer, W. (1986).
A multispecies biofilm model.
\textit{Biotechnology and Bioengineering}, 28(3), 314--328.
\url{https://doi.org/10.1002/bit.260280304}

\bibitem{xavier2005}
Xavier, J.B., Picioreanu, C., van Loosdrecht, M.C.M. (2005).
A framework for multidimensional modelling of activity and structure of multispecies biofilms.
\textit{Environmental Microbiology}, 7(8), 1085--1103.
\url{https://doi.org/10.1111/j.1462-2920.2005.00787.x}

\bibitem{lardon2011}
Lardon, L.A., et al. (2011).
iDynoMiCS: next-generation individual-based modelling of biofilms.
\textit{Environmental Microbiology}, 13(9), 2416--2434.
\url{https://doi.org/10.1111/j.1462-2920.2011.02414.x}

\bibitem{sonner2015}
Sonner, S., Efendiev, M.A., Eberl, H.J. (2015).
On the well-posedness of a mathematical model of quorum-sensing in patchy biofilm communities.
\textit{Mathematical Methods in the Applied Sciences}, 38(3), 3037--3042.
\url{https://doi.org/10.1002/mma.3237}

\bibitem{diele2015}
Diele, F., Marangi, C., Ragni, S. (2015).
Geometric numerical integration in ecological modelling.
\textit{Mathematics and Computers in Simulation}, 110, 40--52.
\url{https://doi.org/10.1016/j.matcom.2014.02.006}

\bibitem{kuramoto1984}
Kuramoto, Y. (1984).
\textit{Chemical Oscillations, Waves, and Turbulence}.
Springer-Verlag, Berlin.
\url{https://doi.org/10.1007/978-3-642-69689-3}

\bibitem{acebron2005}
Aceb\'r\'on, J.A., Bonilla, L.L., P\'erez Vicente, C.J., Ritort, F., Spigler, R. (2005).
The Kuramoto model: a simple paradigm for synchronization phenomena.
\textit{Reviews of Modern Physics}, 77(1), 137--185.
\url{https://doi.org/10.1103/RevModPhys.77.137}

\bibitem{campillo2017}
Campillo, F., Joannides, M., Larramendy-Valverde, I. (2017).
Stochastic analysis of a full system of two competing populations in a chemostat.
\textit{Chemical Engineering Science}, 175, 424--440.
\url{https://doi.org/10.1016/j.ces.2017.10.052}

\bibitem{heidelberg2002}
Heidelberg, J.F., et al. (2002).
Genome sequence of the dissimilatory metal ion-reducing bacterium
\textit{Shewanella oneidensis}.
\textit{Nature Biotechnology}, 20(11), 1118--1123.
\url{https://doi.org/10.1038/nbt749}

\bibitem{veeramani2011}
Veeramani, H., et al. (2011).
Products of abiotic U(VI) reduction by biogenic magnetite and vivianite.
\textit{Geochimica et Cosmochimica Acta}, 75(9), 2512--2528.
\url{https://doi.org/10.1016/j.gca.2011.02.024}

\bibitem{shukla2020}
Shukla, A., Parmar, P., Goswami, D., Patel, B., Saraf, M. (2020).
Characterization of novel thorium tolerant \textit{Ochrobactrum intermedium} AM7.
\textit{Journal of Hazardous Materials}, 388, 122047.
\url{https://doi.org/10.1016/j.jhazmat.2020.122047}

\bibitem{wright1932}
Wright, S. (1932).
The roles of mutation, inbreeding, crossbreeding and selection in evolution.
\textit{Proceedings of the Sixth International Congress of Genetics}, 1, 356--366.

\bibitem{kauffman1989}
Kauffman, S.A., Weinberger, E.D. (1989).
The NK model of rugged fitness landscapes.
\textit{Journal of Theoretical Biology}, 141(2), 211--245.
\url{https://doi.org/10.1016/S0022-5193(89)80019-0}

\bibitem{berg1993}
Berg, H.C. (1993).
\textit{Random Walks in Biology}.
Princeton University Press.

\bibitem{guttenplan2013}
Guttenplan, S.B., Shaw, S., Kearns, D.B. (2013).
The cell biology of peritrichous flagella in \textit{Bacillus subtilis}.
\textit{Molecular Microbiology}, 87(1), 211--229.
\url{https://doi.org/10.1111/mmi.12103}

\bibitem{nicholson2000}
Nicholson, W.L., et al. (2000).
Resistance of \textit{Bacillus} endospores to extreme terrestrial and extraterrestrial
environments.
\textit{Microbiology and Molecular Biology Reviews}, 64(3), 548--572.
\url{https://doi.org/10.1128/MMBR.64.3.548-572.2000}

\bibitem{turick2011}
Turick, C.E., et al. (2011).
Gamma radiation interacts with melanin to alter its oxidation-reduction potential and
results in electric current production.
\textit{Bioelectrochemistry}, 82(1), 69--73.
\url{https://doi.org/10.1016/j.bioelechem.2011.04.009}

\bibitem{robertson2012}
Robertson, K.L., et al. (2012).
Adaptation of the black yeast \textit{Wangiella dermatitidis} to ionizing radiation.
\textit{PLoS ONE}, 7(11), e48674.
\url{https://doi.org/10.1371/journal.pone.0048674}

\bibitem{malo2018}
Malo, M.E., et al. (2018).
Morphological changes in melanized and non-melanized \textit{Cryptococcus neoformans} cells
post exposure to ionizing radiation.
\textit{Fungal Biology}, 122(6), 449--456.
\url{https://doi.org/10.1016/j.funbio.2017.08.012}

\bibitem{newsome2014}
Newsome, L., Morris, K., Lloyd, J.R. (2014).
The biogeochemistry and bioremediation of uranium and other priority radionuclides.
\textit{Chemical Geology}, 363, 164--184.
\url{https://doi.org/10.1016/j.chemgeo.2013.10.034}

\bibitem{brim2000}
Brim, H., et al. (2000).
Engineering \textit{Deinococcus radiodurans} for metal remediation in radioactive mixed
waste environments.
\textit{Nature Biotechnology}, 18(1), 85--90.
\url{https://doi.org/10.1038/71986}

\bibitem{kazy2009}
Kazy, S.K., D'Souza, S.F., Sar, P. (2009).
Uranium and thorium sequestration by a \textit{Pseudomonas} sp.
\textit{Journal of Hazardous Materials}, 163(1), 65--72.
\url{https://doi.org/10.1016/j.jhazmat.2008.06.076}

\bibitem{blasius1999}
Blasius, B., Huppert, A., Stone, L. (1999).
Complex dynamics and phase synchronization in spatially extended ecological systems.
\textit{Nature}, 399, 354--359.
\url{https://doi.org/10.1038/20676}

\bibitem{alpkvist2007}
Alpkvist, E., Klapper, I. (2007).
A multidimensional multispecies continuum model for heterogeneous biofilm development.
\textit{Bulletin of Mathematical Biology}, 69(2), 765--789.
\url{https://doi.org/10.1007/s11538-006-9168-7}

\bibitem{eberl2001}
Eberl, H.J., Parker, D.F., van Loosdrecht, M.C.M. (2001).
A new deterministic spatio-temporal continuum model for biofilm development.
\textit{Journal of Theoretical Medicine}, 3(3), 161--175.
\url{https://doi.org/10.1080/10273660108833072}

\bibitem{hairer2006}
Hairer, E., Lubich, C., Wanner, G. (2006).
\textit{Geometric Numerical Integration}, 2nd ed.
Springer-Verlag.
\url{https://doi.org/10.1007/3-540-30666-8}

\bibitem{eisenman2012}
Eisenman, H.C., Casadevall, A. (2012).
Synthesis and assembly of fungal melanin.
\textit{Applied Microbiology and Biotechnology}, 93(3), 931--940.
\url{https://doi.org/10.1007/s00253-011-3777-2}

\bibitem{lloyd2005}
Lloyd, J.R., Renshaw, J.C. (2005).
Bioremediation of radioactive waste.
\textit{Current Opinion in Biotechnology}, 16(3), 254--260.
\url{https://doi.org/10.1016/j.copbio.2005.04.012}

\bibitem{khajo2011}
Khajo, A., et al. (2011).
Protection of melanized \textit{Cryptococcus neoformans} from lethal dose gamma irradiation.
\textit{PLoS ONE}, 6(9), e25092.
\url{https://doi.org/10.1371/journal.pone.0025092}

\bibitem{battista1997}
Battista, J.R. (1997).
Against all odds: the survival strategies of \textit{Deinococcus radiodurans}.
\textit{Annual Review of Microbiology}, 51, 203--224.
\url{https://doi.org/10.1146/annurev.micro.51.1.203}

\bibitem{casadevall2017}
Casadevall, A., et al. (2017).
Melanin, radiation, and energy transduction in fungi.
\textit{Microbiology Spectrum}, 5(2), FUNK-0037-2016.
\url{https://doi.org/10.1128/microbiolspec.FUNK-0037-2016}

\end{thebibliography}

\end{document}
